\chapter{Specifications For Dissertation Format}

\section{Preparation of Manuscript and Copies }
The dissertation needs to be prepared using a standard text processing software and should be printed in black text (color for images, if necessary) using a laser printer or letter quality printer in standard typeface (Times New Roman).\\
The dissertation must be printed or photocopied on single sides of good quality white paper \cite{atzori2010}.\\

All copies of dissertation pages must be clear, sharp and even, with uniform size and uniformly spaced characters, lines and margins on every page of good quality white paper of 75 gsm or more\cite{bakshicisco}.\\
 
Dissertation should be free from typographical errors.\\

The Chapter Header should be of size 18, Typeface: Times New Roman with Bold \& Underline Tag, Header/Title should be of size 14, Typeface: Times New Roman and Bold, Sub-section/Text\cite{flexisecjinwala2012}.\\ 

Content should be of size 12, Typeface: Times New Roman.

\section{Size and Margins}
\subsection{Printing of Dissertation}
A4 (21 cm x 29.7 cm) is the recommended dissertation size. Dissertation should be printed SINGLE SIDE\cite{tinyseckarlof2004}. 
\subsection{}
The top and bottom margins should be 1.5 and 1 cm respectively, whereas, the left side (inside) and right side (outside) margin should be 2.5 and 1 cm respectively for both textual and non-textual (e.g., figures, tables) pages. 
\subsection{}
Content should not extend beyond the bottom margin except for completing a footnote, last line of chapter/subdivision, or figure/table caption.
\subsection{}
A sub-head at the bottom of the page should have at least two full lines of content below it. If the sub-head is too short to allow this, it should begin on the next page. 
\subsection{}
All tables and figures should conform to the same requirements as text. Color may be used for figures. If tables and figures are large, they may be reduced to the standard size (provided the reduced area is not less than 50\% of the original) and/or folded just once to flush with the dissertation margin. 
\subsection{}
Students should also submit the dissertation in soft form (both PDF and word) for storage and archival

\section{Page Numbering}
\subsection{}
Beginning with the first page of the text in the dissertation (chapter 1), all pages should be numbered consecutively and consistently in Arabic numerals (1, 2, 3, …) through the appendices. All these pages should be Single page printed.
\subsection{}
Page numbers prior to Chapter 1 should be in lower case Roman numerals (i, ii, iii, …). The title page is considered to be page (i) but the number is not printed. All these pages should be single page printed. 

\subsection{}
All page numbers should be placed without punctuation in the lower right-hand corner, 1. 2 cm from the top edge and with the last digit even with the right-hand margin.

\section{Line Spacing}
The general text of the manuscript should be in 1.5 spacing (3 lines per inch). Long tables, quotations, footnotes, multi-line captions and bibliographic entries (references) should be in single spacing (6 lines per inch), with a preferred text size in 12 points. 

